From noise to knowledge: how randomness generates novel phenomena and reveals information

### 2019-10-17 paper/paper.Rmd
fix typo

Applying the system size approximation to the example of the Levins model introduced
previously lets us both recover the classic model while adding some nuance in 
the interpretation.
Taking $b(n) = n (1 - \tfrac{n}{N})$ and $d(n) = en$ and solving for the steady
state to van Kampen's ODEs, we see the average number $\langle n \rangle$  of occupied patches is
$\langle n \rangle = N\tfrac{e}{c}$ with variance at steady-state of 
$\sigma^2_n = N \tfrac{e}{c}$.  Observe that the variance scales proportionally
to the system size, $N$, which explains the standard deviation
increases as $\sqrt{N}$, as in the canonical model and as predicted by Figure \ref{gillespie}.
We can also see that different ways of partitioning the same macroscopic equation
into components of birth and death will create the same mean dynamics
but with *different* levels of noise.  For instance, in a logistic
model with $b = rn$ and $d = rn^2/N$, we instead find the variance
at steady state is $\sigma^2 = K$; significantly larger than in the
Levins model.

>>>>

Applying the system size approximation to the example of the Levins model introduced
previously lets us both recover the classic model while adding some nuance in 
the interpretation.
Taking $b(n) = n (1 - \tfrac{n}{N})$ and $d(n) = en$ and solving for the steady
state to van Kampen's ODEs, we see the average number $\langle n \rangle$  of occupied patches is
$\langle n \rangle = N\left(1 - \tfrac{e}{c}\right)$ with variance at steady-state of 
$\sigma^2_n = N \tfrac{e}{c}$.  Observe that the variance scales proportionally
to the system size, $N$, which explains the standard deviation
increases as $\sqrt{N}$, as in the canonical model and as predicted by Figure \ref{gillespie}.
We can also see that different ways of partitioning the same macroscopic equation
into components of birth and death will create the same mean dynamics
but with *different* levels of noise.  For instance, in a logistic
model with $b = rn$ and $d = rn^2/N$, we instead find the variance
at steady state is $\sigma^2 = K$; significantly larger than in the
Levins model.
